### 2025-07-24 conteudo/background/codeduplication.tex
another grammar typos fix

Code Duplication Detection is a field in computer science that has attracted
researchers' attention since 1988 \citep{firstman}. The presence of code
duplication is considered a harmful artifact in software, as it negatively
affects tasks such as code readability and increases the likelihood of
introducing bugs \citep{harmone}. In most cases, code duplication tends to make
software more unstable compared to its non-duplicated counterpart
\citep{harmtwo}.

@@

The classification of two code artifacts as duplicates raises the question of
what constitutes duplication. To address this, the literature categorizes code
duplication based on the extent of differences between code artifacts,
resulting in four types of code clone duplications \citep{litreview}. When
comparing two code artifacts, F and G, the four types of code duplication can
be described as follows \citep{litreview}:

@@

Using the counts of TP, FP, TN, and FN, standard metrics such as Precision,
Recall, and F-Score can be calculated \citep{recall}. In this work, we focus on
the Recall metric, as it is commonly reported in literature for comparing code
clone detection approaches. Recall is defined as follows \citep{recall}:

>>>>

Code Duplication Detection is a field in computer science that has attracted
researchers' attention since 1988 \citep{firstman}. Code
duplication is considered a harmful artifact in software, as it negatively
affects tasks such as code readability and increases the likelihood of
introducing bugs \citep{harmone}. In most cases, code duplication tends to make
software more unstable than its non-duplicated counterpart
\citep{harmtwo}.

@@

Classifying two code artifacts as duplicates raises the question of
what constitutes duplication. To address this, the literature categorizes code
duplication based on the extent of differences between code artifacts,
resulting in four types of code clone duplications \citep{litreview}. When
comparing two code artifacts, F and G, the four types of code duplication can
be described as follows \citep{litreview}:

@@

Using the counts of TP, FP, TN, and FN, standard metrics such as Precision,
Recall, and F-Score can be calculated \citep{recall}. In this work, we focus on
the Recall metric, which is commonly reported in literature for comparing code
clone detection approaches. Recall is defined as follows \citep{recall}:

### 2025-07-24 conteudo/background/linux.tex
add lines from IIO
Update typo on IIO metrics
another grammar typos fix

This chapter presents an overview of fundamental topics needed to understand 
the research, such as the Linux Kernel, refactoring, types of code duplication, 
code duplication detection methods presented in the literature and ways to 
evaluate these methods. This foundation will make it easier to grasp the 
concepts discussed in the research.

@@

The Linux kernel, a Unix-like operating system kernel, was published by Linus
Torvalds in 1992 as free software. It began gaining popularity in the late
1990s \citep{linuxbook}. Contrary to popular belief, Linux itself is not a
complete operating system, as it lacks components like filesystem utilities and
windowing systems. The so-called ``Linux distributions'' that are commonly
recognized by the public are actually GNU/Linux operating systems, which
include various essential applications alongside the Linux kernel,
such as system
\citep{gnuref}. Examples of Linux distributions are Ubuntu \citep{ubuntu} 
and Arch Linux \citep{archlinux}.

@@

Mailing lists are the primary means of communication within the Linux
community. Members of the community use them to submit and review patches,
discuss Linux-related topics, and debate potential new features. Given the
kernel's extensive modularity, each subsystem has its mailing list, allowing
discussions to be organized around specific areas of development
\citep{melissa}. All interactions on these mailing lists are documented and
archived in the Linux Kernel Lore \citep{linuxlore}, which serves as a
repository for kernel discussions and can be accessed by anyone interested.

One of the subsystems within the Linux kernel is the AMD Display driver, which
is responsible for implementing the drivers needed for AMD GPUs to function
correctly in a Linux environment. The maintainers of this subsystem have
reported a practical issue: a significant amount of duplicated code, which
complicates the maintenance of this subsystem. This problem serves as the
starting point for our research, with the AMD Display driver as our primary
object of study as we aim to understand and mitigate code duplication in the
Linux kernel. The AMD Display driver subsystem contains more then 391000 lines 
of code and more then 1089 code files
\footnote{ 
Data extracted directly from the Linux repository with the help of the cloc tool.
\citep{cloc}
}, 
making it a considerable component within the Linux kernel. Like other 
subsystems, discussions about the AMD Display driver are conducted on its 
designated mailing list, which can be accessed here: 
\url{https://lists.freedesktop.org/mailman/listinfo/amd-gfx}.

Another subsystem within the Linux kernel explored in this work is the
Industrial I/O subsystem (IIO). According to the Linux kernel documentation, 
''The main purpose of the Industrial I/O subsystem (IIO) is to provide support 
for devices that in some sense perform either analog-to-digital conversion (ADC) 
or digital-to-analog conversion (DAC) or both''\citep{iiodoc}. The IIO subsystem 
contains more than X lines of code and more than Y code files
\footnote{
Data extracted directly from the Linux repository with the help of the clol tool. \citep{cloc}
}.
We opted to include the IIO subsystem in our experiments to ensure our findings 
are not limited to one subsystem. Additionally, we have an established point of
contact within the IIO subsystem to facilitate this work.
Like other subsystems, discussions about the IIO subsytem are conducted on its 
designated mailing list, which can be found here: 
\url{https://subspace.kernel.org/vger.kernel.org.html}.

@@

We use the \textit{diff} command as the base of one of the code duplication detection 
methods on the tool proposed in this work, although it is not explored as the text 
similarity method for code duplication detection.

>>>>

This chapter presents an overview of fundamental topics needed to understand 
the research, such as the Linux Kernel, refactoring, types of code duplication, 
code duplication detection methods presented in the literature, and ways to 
evaluate these methods. This foundation will make it easier to grasp the 
concepts discussed in the research.

@@

The Linux kernel, a Unix-like operating system kernel, was published by Linus
Torvalds in 1992 as free software. It began gaining popularity in the late
1990s \citep{linuxbook}. Contrary to popular belief, Linux is not a
complete operating system, as it lacks components like filesystem utilities and
windowing systems. The so-called ``Linux distributions'' commonly
recognized by the public are GNU/Linux operating systems, which
include various essential applications alongside the Linux kernel,
such as system
\citep{gnuref}. Examples of Linux distributions are Ubuntu \citep{ubuntu} 
and Arch Linux \citep{archlinux}.

@@

Mailing lists are the primary means of communication within the Linux
community. Community members use them to submit and review patches,
discuss Linux-related topics, and debate potential new features. Given the
kernel's extensive modularity, each subsystem has its mailing list, allowing
discussions to be organized around specific areas of development
\citep{melissa}. All interactions on these mailing lists are documented and
archived in the Linux Kernel Lore \citep{linuxlore}, which serves as a
repository for kernel discussions and can be accessed by anyone interested.

One of the subsystems within the Linux kernel is the AMD Display driver, which
is responsible for implementing the drivers needed for AMD GPUs to function
correctly in a Linux environment. The maintainers of this subsystem have
reported a practical issue: a significant amount of duplicated code, which
complicates the maintenance of this subsystem. This problem serves as the
starting point for our research, with the AMD Display driver as our primary
object of study as we aim to understand and mitigate code duplication in the
Linux kernel. The AMD Display driver subsystem contains more than 391000 lines 
of code and more than 1089 code files
\footnote{ 
Data extracted directly from the Linux repository with the help of the cloc tool.
\citep{cloc}
}, 
making it a considerable component within the Linux kernel. Like other 
subsystems, discussions about the AMD Display driver are conducted on its 
designated mailing list, which can be accessed here: 
\url{https://lists.freedesktop.org/mailman/listinfo/amd-gfx}.

Another subsystem within the Linux kernel explored in this work is the
Industrial I/O subsystem (IIO). According to the Linux kernel documentation, 
''The main purpose of the Industrial I/O subsystem (IIO) is to provide support 
for devices that in some sense perform either analog-to-digital conversion (ADC) 
or digital-to-analog conversion (DAC) or both''\citep{iiodoc}. The IIO subsystem 
contains more than 281772 lines of code and more than 755 code files
\footnote{
Data extracted directly from the Linux repository with the help of the clol tool. \citep{cloc}
}.
We opted to include the IIO subsystem in our experiments to ensure our findings 
are not limited to one subsystem. Additionally, we have an established point of
contact within the IIO subsystem to facilitate this work.
Like other subsystems, discussions about the IIO subsystem are conducted on its 
designated mailing list, which can be found here: 
\url{https://subspace.kernel.org/vger.kernel.org.html}.

@@

We use the \textit{diff} command as the base of one of the code duplication detection 
methods on the tool proposed in this work, although it is not explored as the text 
similarity method chosen for code duplication detection.

### 2025-07-24 conteudo/background/nlpsimilarity.tex
another grammar typos fix

The standard methodology for text similarity detection is based on building
vector embeddings of texts and applying a distance function to these embeddings
to measure the similarity between them. The literature proposes multiple vector
embedding methods, such as bag-of-words, TF-IDF, LSI, and Word2Vec
\citep{gensimlivro}. Recent works related to vector embedding construction
utilize Large Language Models (LLMs), as exemplified by Ethayarajh's work,
which employs BERT, ELMo, and GPT \citep{llmsimilar}. The most popular distance
function used in the literature is cosine similarity, defined by the formula
\citep{cosineref}:

@@

where $A$ and $B$ are the vector embeddings of the two texts. An advantage of
cosine similarity is that $SC(A,B)$ produces a number between $0$ and $1$,
where values closer to $1$ indicate higher similarity. It is common to
interpret cosine similarity as an approximation of probability, transforming
values between $0$ and $1$ into percentages between $0\%$ and $100\%$.
Throughout our research, we will use this percentage form.

It is well-known that LLMs have a high computational cost. For instance, an
analysis by Reimers and Gurevych found that finding the most similar pair in a
collection of 10,000 sentences required about 65 hours with BERT
\citep{bertsimilar}. Since we aim to compare a similar magnitude of pairs in
our work, we decided not to use LLMs for our vector embedding method, even
though they may be suitable for smaller codebases.

>>>>

The standard methodology for text similarity detection is based on building
vector embeddings of texts and applying a distance function to these embeddings
to measure their similarity. The literature proposes multiple vector
embedding methods, such as bag-of-words, TF-IDF, LSI, and Word2Vec
\citep{gensimlivro}. Recent works related to vector embedding construction
utilize Large Language Models (LLMs), as exemplified by Ethayarajh's work,
which employs BERT, ELMo, and GPT \citep{llmsimilar}. The most popular distance
function used in the literature is cosine similarity, defined by the formula
\citep{cosineref}:

@@

Where $A$ and $B$ are the vector embeddings of the two texts. An advantage of
cosine similarity is that $SC(A,B)$ produces a number between $0$ and $1$,
where values closer to $1$ indicate higher similarity. It is common to
interpret cosine similarity as an approximation of probability, transforming
values between $0$ and $1$ into percentages between $0\%$ and $100\%$.
Throughout our research, we will use this percentage form.

It is well-known that LLMs have a high computational cost. For instance, an
analysis by Reimers and Gurevych found that finding the most similar pair in a
collection of 10,000 sentences required about 65 hours with BERT
\citep{bertsimilar}. Since we aim to compare pairs of similar magnitude in
our work, we decided not to use LLMs for our vector embedding method, even
though they may be suitable for smaller codebases.

### 2025-07-24 conteudo/conclusion/conclusion.tex
another grammar typos fix

In response to these limitations, this research addressed the challenge of 
code duplication within the Linux Kernel through the development of ArKanjo, 
command-line tool for Linux kernel maintenance designed to detect and 
analyze function-level duplications. Released under the MIT license, 
the architectural design of ArKanjo, incorporating a Preprocessor and a 
Query Responder, separates computationally intensive analysis from efficient 
querying for duplications within large codebases. 
Validation procedures, employing both the BigCloneBench dataset and empirical 
analysis conducted on the AMD Display driver, indicated ArKanjo's proficient 
performance in identifying Type-1 and Type-2 duplications, notably when 
similarity thresholds are configured between 80\% and 100\%.

The ethnographic studies, which involved participant observation and a project 
with university students, provided a realistic view of both the opportunities 
and the considerable challenges in mitigating code duplications found by ArKanjo. 
The tool was effectively used by first-time contributors to the Linux kernel to 
identify duplications and prepare patches. While this research included a
successfully merged patch by the author, and successful submissions from many 
student groups, a notable portion of student efforts 
encountered significant hurdles. These difficulties often stemmed from maintainer 
feedback where proposed changes, though reducing duplication, were rejected or 
required substantial rework due to concerns about code readability, added abstraction, 
or the specific context of the code. Technical complexities, such as C macros in 
configuration files, and lengthy patch review times also contributed to the challenges. 
Furthermore, the studies confirmed that maintainers do not always view code duplication 
as inherently negative, especially when it may improve clarity or facilitate 
independent hardware support.

In its entirety, this research contributes the ArKanjo tool as an operational 
method for detecting code duplication, and importantly, provides insights into 
the process of managing such duplications within the Linux kernel subsystems 
studied. The results show that while ArKanjo can effectively identify 
duplicated functions, the path from detection to an accepted mitigation within 
the kernel involves careful consideration of the specific code, established 
community practices, and the maintainers' perspectives on code quality and readability.

@@

The work presented in this thesis suggests several directions for future investigation. 
Enhancements to the ArKanjo tool could be pursued, such as investigating a multi-stage 
detection strategy. This strategy might employ a computationally lighter method for an 
initial, broad pass to identify candidate duplications, followed by more 
resource-intensive, state-of-the-art techniques for refined analysis and improved 
detection of complex clone types. Such an approach could complement efforts for 
improved resource management during its preprocessing phase for very large codebases.
The tool's language support could also be broadened by
developing parsers for additional programming languages, building upon its adaptable
architecture. Beyond the tool itself, future research could extend the empirical
analysis of code duplication and mitigation strategies to a broader range of Linux kernel 
subsystems or other large-scale C projects, to determine the wider 
applicability of the observations made.

>>>>

In response to these limitations, this research addressed the challenge of code 
duplication within the Linux Kernel by developing ArKanjo, a command-line tool 
for Linux kernel maintenance designed to detect and analyze function-level duplications. 
Released under the MIT license, the architectural design of ArKanjo, incorporating a 
Preprocessor and a Query Responder, separates computationally intensive analysis from 
efficient querying for duplications within large codebases. Validation procedures, 
employing both the BigCloneBench dataset and empirical analysis conducted on the AMD 
Display driver, indicated ArKanjo's proficient performance in identifying Type-1 and 
Type-2 duplications, notably when similarity thresholds are configured between 80\% and 100\%.

The ethnographic studies, which involved participant observation and a project with university 
students, provided a realistic view of the opportunities and the considerable challenges in 
mitigating code duplications found by ArKanjo. The tool was effectively used by first-time 
contributors to the Linux kernel to 
identify duplications and prepare patches. While this research included a successfully merged 
patch by the author, and successful submissions from many student groups, a notable portion of 
student efforts encountered significant hurdles. These difficulties often stemmed from maintainer 
feedback where proposed changes, though reducing duplication, were rejected or required 
substantial rework due to concerns about code readability, added abstraction, or the specific 
context of the code. Technical complexities, such as C macros in configuration files, and lengthy 
patch review times also contributed to the challenges. Furthermore, the studies confirmed that 
maintainers do not always view code duplication as inherently harmful, especially when it may 
improve clarity or facilitate independent hardware support.

This research contributes to the ArKanjo tool as an operational method for detecting code 
duplication and, importantly, provides insights into the process of managing such duplications 
within the Linux kernel subsystems studied. The results show that while ArKanjo can effectively 
identify duplicated functions, the path from detection to an accepted mitigation within the kernel 
involves careful consideration of the specific code, established community practices, and the 
maintainers' perspectives on code quality and readability.

@@

The work presented in this thesis suggests several directions for future investigation. Enhancements 
to the ArKanjo tool could be pursued, such as investigating a multi-stage detection strategy. 
This strategy might employ a computationally lighter method for an initial, broad pass to identify 
candidate duplications, followed by more resource-intensive, state-of-the-art techniques for refined 
analysis and improved detection of complex clone types. Such an approach could complement efforts for 
improved resource management during its preprocessing phase for large codebases. The tool's language 
support could be broadened by developing parsers for additional programming languages, building upon 
its adaptable architecture. Beyond the tool itself, future research could extend the empirical analysis 
of code duplication and mitigation strategies to a broader range of Linux kernel subsystems or other 
large-scale C projects, to determine the broader applicability of the observations made.

### 2025-07-24 conteudo/introduction/intro.tex
another grammar typos fix

The Linux kernel is a widely used Free/Libre/Open Source Software (FLOSS) project, 
powering 96.3\% of the top one million web servers and 85\% of smartphones worldwide. 
Maintaining the kernel is an enormous task, involving more than 28 million lines of 
code and contributions from over twenty thousand developers
\footnote{Data directly extracted from Linux github repository \citep{linuxrepo}.
Counted  28783390 lines of code and 20699 number of contributors in May 29th, 2025.
Number of contributors counted as number of unique full name listed as author of a patch.}.
Due to the scale of the project, not all contributions adhere to best practices, resulting 
in poor-quality code artifacts that may complicate maintenance and future feature development.

One such artifact is code duplication, which is the focus of this research. The motivation for 
this work began with a practical problem rooted in device drivers, which are a major part 
of the kernel, representing 66\% of the source code\citep{marcelo}. We interacted with the 
maintainers of the AMD Display driver, an essential component for the 19\% of personal GPUs 
manufactured by AMD in 2023\citep{gpumarket}. They shared their challenges with a significant 
amount of duplicated code that hampers the driver's maintenance. In searching for a solution, 
we found that existing tools are not well-suited for identifying duplications in large-scale 
codebases like the Linux kernel, nor do they offer guidance on how to resolve the duplications 
they find. Despite searching both formal and grey literature, existing solutions fell short of 
addressing these specific challenges. The list of related tools examined in this exploratory 
research is available in Appendix \ref{app:gray}.

@@

To address the code duplication problem within the Linux kernel, this research employs a 
multi-method approach designed to bridge the gap between detection and practical mitigation. 
Our approach, illustrated in Figure \ref{fig:reDesign}, is structured into two interconnected phases that progress 
from tool development to an in-depth study of its application within the kernel’s development community. 
The overall objective is to produce not only a functional tool but also actionable insights into 
the patterns and practices surrounding code duplication and its resolution.

@@

The initial phase of our research focused on the development and validation of ArKanjo, a 
command-line tool capable of identifying code duplication at the function level.  The 
tool's accuracy was rigorously validated by evaluating its performance against the 
standard BigCloneBench dataset \citep{bigclonebench} and by conducting an empirical analysis 
of its results within the AMD Display driver.  This foundational work ensured we had a 
reliable instrument to proceed with the practical investigation of code duplication 
in our target subsystems.

With a validated tool, the research then moved from detection to practice, investigating 
the real-world challenges of mitigating the identified duplications. This was accomplished 
through an ethnographic study designed to engage with the kernel community and understand 
their perspectives on code quality and refactoring.  This study had two components: a 
participant-observation experiment where the author submitted refactoring patches to the 
AMD Display driver, and a non-participant observation of university students using ArKanjo 
to contribute fixes to both the AMD Display and Industrial I/O subsystems.

>>>>

The Linux kernel is a widely used Free/Libre/Open Source Software (FLOSS) project, 
powering 96.3\% of the top one million web servers and 85\% of smartphones worldwide. 
Maintaining the kernel is an enormous task, involving more than 28 million lines of 
code and contributions from over twenty thousand developers
\footnote{Data directly extracted from Linux github repository \citep{linuxrepo}.
Counted  28783390 lines of code and 20699 contributors on May 29th, 2025.
Number of contributors counted as unique full names listed as a patch's author.}.
Due to the scale of the project, not all contributions adhere to best practices, resulting 
in poor-quality code artifacts that may complicate maintenance and future feature development.

One such artifact is code duplication, which is the focus of this research. The motivation for 
this work began with a practical problem rooted in device drivers, which are a significant part 
of the kernel, representing 66\% of the source code\citep{marcelo}. We interacted with the 
maintainers of the AMD Display driver, an essential component for the 19\% of personal GPUs 
manufactured by AMD in 2023\citep{gpumarket}. They shared their challenges with a significant 
amount of duplicated code that hampers the driver's maintenance. In searching for a solution, 
we found that existing tools are not well-suited for identifying duplications in large-scale 
codebases like the Linux kernel, nor do they offer guidance on resolving the duplications 
they find. Despite searching formal and grey literature, existing solutions failed to address
addressing these specific challenges. The list of related tools examined in this exploratory 
research is available in Appendix \ref{app:gray}.

@@

To address the code duplication problem within the Linux kernel, this research employs a 
multi-method approach to bridge the gap between detection and practical mitigation. 
Our approach, illustrated in Figure \ref{fig:reDesign}, is structured into two interconnected phases that progress 
from tool development to an in-depth study of its application within the kernel’s development community. 
The overall objective is to produce a functional tool and actionable insights into 
the patterns and practices surrounding code duplication and its resolution.

@@

The initial phase of our research focused on developing and validating ArKanjo, a
command-line tool capable of identifying code duplication at the function level. 
We rigorously validated the tool's accuracy by evaluating its performance against the
standard BigCloneBench dataset \citep{bigclonebench} and by conducting an empirical analysis
of its results within the AMD Display driver. This foundational work ensured we had a
reliable instrument to proceed with the practical investigation of code duplication
in our target subsystems.

With a validated tool, the research moved from detection to practice, investigating the 
real-world challenges of mitigating the identified duplications. This was accomplished 
through an ethnographic study to engage with the kernel community and understand their 
perspectives on code quality and refactoring. This study had two components: a 
participant-observation experiment where the author submitted refactoring patches to the 
AMD Display driver, and a non-participant observation of university students using ArKanjo
to contribute fixes to both the AMD Display and Industrial I/O subsystems.

### 2025-07-24 conteudo/methods/methods.tex
another grammar typos fix

We employ a multi-method research approach, using different methodologies to achieve our 
goals to approach code duplication in some of the Linux subsystems. To validate if the 
ArKanjo tool can find code duplications, we applied two independent methods and performed 
triangulation on the results.

The first method evaluates the proposed tool using an approach from the literature, 
comparing it against the BigCloneBench dataset \citep{bigclonebench}. 
The second method is an empirical analysis of a subset of functions within the 
AMD Display driver that our tool identified as duplicates. After completing these two 
methods, we triangulated the findings to analyze the results comprehensively. 
We chose triangulation because relying solely on the literature-based method has 
limitations and may be unreliable \citep{bigfail, litreview}. 
As a reminder, all research approaches were done using the text similarity method 
as the code duplication detection method of the tool.

@@

The ethnographic study had two phases. In the first phase, the author explored the process 
of mitigating a code duplication found in the AMD Display driver and sent it as a PATCH to 
interact with the Linux community and get their feedback. In the second phase, we proposed 
to students in the course Free Software Development at the University of São Paulo to mitigate 
the code duplications found by the tool in the IIO subsystem and the AMD Display driver and 
pass the process of sending the mitigations as PATCH.

@@

Due to the limitations of the BigCloneBench dataset, as noted by \cite{bigfail}, and its lack of suitability 
for our context (Java dataset vs. C-based Linux kernel), we complemented the BigCloneBench evaluation with 
an empirical study of the tool inferred duplications.

@@

This section outlines our approach to mitigating code duplication in the Linux kernel 
and describes the ethnographic methodology for engaging with the Linux community. 
Firstly, we present the participant observation research method explored by the author. 
Then, we present the methodology of the non-participant observation analyze of patches 
sent by the students.

>>>>

We employ a multi-method research approach, using different methodologies to achieve our 
goals of approaching code duplication in some of the Linux subsystems. To validate if the 
ArKanjo tool can find code duplications, we applied two independent methods and performed 
triangulation on the results.

The first method evaluates the proposed tool using an approach from the literature, 
comparing it against the BigCloneBench dataset \citep{bigclonebench}. 
The second method is an empirical analysis of a subset of functions within the 
AMD Display driver that our tool identified as duplicates. After completing these two 
methods, we triangulated the findings to analyze the results comprehensively. 
We chose triangulation because relying solely on the literature-based method has 
limitations and may be unreliable \citep{bigfail, litreview}. 
As a reminder, we have done all research approaches using the text similarity method 
as the code duplication detection method of the tool.

@@

The ethnographic study had two phases. In the first phase, the author explored the process 
of mitigating a code duplication found in the AMD Display driver and sent it as a PATCH to 
interact with the Linux community and get their feedback. In the second phase, we proposed 
to students in the course Free Software Development at the University of São Paulo to mitigate 
the code duplications found by the tool in the IIO subsystem and the AMD Display driver and 
pass the process of sending the mitigations as a PATCH.

@@

Due to the limitations of the BigCloneBench dataset, as noted by \cite{bigfail}, and its lack of suitability 
for our context (Java dataset vs. C-based Linux kernel), we complemented the BigCloneBench evaluation with 
an empirical study of the tool's inferred duplications.

@@

This section outlines our approach to mitigating code duplication in the Linux kernel 
and describes the ethnographic methodology for engaging with the Linux community. 
Firstly, we present the participant observation research method explored by the author. 
Then, we present the methodology of the non-participant observation analysis of patches 
sent by the students.

### 2025-07-24 conteudo/results/eval.tex
another grammar typos fix

As described in Section \ref{sec:metbig}, for evaluating our tool on the BigCloneBench dataset \citep{bigclonebench}, we define that every pair of functions with a similarity metric equal to or greater than a threshold \( X \) is considered a duplication. The pair of code files containing these functions are then labeled as duplications. If our tool correctly identifies a code duplication pair, it is counted as correct for the respective code duplication type. If it infers a duplication pair incorrectly, it is counted as incorrect for all code duplication types. Table \ref{tab:bigclone} shows the results obtained by our tool on the BigCloneBench dataset \citep{bigclonebench}.

@@

The results in Table \ref{tab:emp} show that as the similarity metric decreases, 
there is a notable increase in function pairs that are not duplications within 
the Linux kernel. For similarity ranges of 79\%-81\% or higher, there is a high 
percentage of code duplications. At 69\%-71\% similarity, the duplication rate 
drops to 50\%. At 59\%-61\%, it further decreases to 20\%, and below 49\%-51\%, 
no code duplications were found.

>>>>

As described in Section \ref{sec:metbig}, for evaluating our tool on the BigCloneBench dataset \citep{bigclonebench}, we define 
that every pair of functions with a similarity metric equal to or greater than a threshold \( X \) is considered a duplication. 
The pair of code files containing these functions is then labeled as duplications. If our tool correctly 
identifies a code duplication pair, it is counted as correct for the respective code duplication type. If it infers a 
duplication pair incorrectly, it is counted as incorrect for all code duplication types. Table \ref{tab:bigclone} shows the results obtained by our tool on the BigCloneBench dataset \citep{bigclonebench}.

@@

The results in Table \ref{tab:emp} show that as the similarity metric decreases, 
there is a notable increase in function pairs that are not duplications within 
the context of the AMD Display driver (the Linux kernel component studied here). 
For similarity ranges of 79\%-81\% or higher, there is a high 
percentage of code duplications. At 69\%-71\% similarity, the duplication rate 
drops to 50\%. At 59\%-61\%, it further decreases to 20\%, and below 49\%-51\%, 
no code duplications were found.

### 2025-07-24 conteudo/results/students.tex
remove todo as I need to submit the text
another grammar typos fix

As presented in Chapter \ref{cha:method}. It was proposed to students, as one of the activities 
of the Free Software course, that they make the first-time contribution to the Linux kernel to 
mitigate duplication found by the ArKanjo tool while sending patches to the Linux subsystem 
maintainers and documenting their experiences and learnings.

@@

\todo[inline]{Is it ok to add it here in the text? It was Spessoto that writed this section entirely and I reviewed it}

@@

We asked the teaching assistant and the students who used the ArKanjo to answer a survey with 
some questions related to the tool experience, such as installation, documentation, facility of 
use, new ideas to improve the tool, etc. Appendix \ref{app:survey} has the complete survey 
with the answers.

@@

Related to the tool's functionalities, both the student assistant and the student reported 
a missing mechanism to save the tool results on a separate file instead of the terminal output. 
The tool, by default, outputted with colors, as shown in section \ref{subsec:func}, which complicates 
redirecting the terminal output to a file.

We asked for the overall impression of the tool and if it met their expectations. We also asked 
them to give a score of one to five for installation, documentation, preprocessing, and functionalities. 
Both the teaching assistant and the student scored at least four on each mentioned point, thought that 
the tool fulfilled their expectations, and had an overall good experience. Finally, the teaching assistant 
the the student reported features ideas to the tool that may be interesting, which 
are showed below.

@@

With the survey feedback, we understood that users had a good experience and found value in the 
ArKanjo tool. They also provided critical points that could be improved. The most critical 
reported issues, such as documentation errors and a lack of support for multiple Linux distributions 
and different operating systems, have been fixed. Less critical issues remain open points that 
can be addressed in future work.

@@

The second group opted to work on a webpage for the tool documentation and support for 
light and dark modes. The colored output presented in the figures in Chapter \ref{cha:tool} was 
hard coded at first, not considering the user terminal colors, thus giving a bad user experience. 
The website helps facilitate people's contribution or use of the tool. We did not know how to 
approach the light and dark modes at first and did not need to look at how to fix the issue as 
the group searched and found a solution.

>>>>

As presented in Chapter \ref{cha:method}. It was proposed to students, as one of the activities 
of the Free Software course, they make their first contribution to the Linux kernel to 
mitigate duplication found by the ArKanjo tool while sending patches to the Linux subsystem 
maintainers and documenting their experiences and learnings.

@@

@@

We asked the teaching assistant and the students who used the ArKanjo to answer a survey with 
some questions related to the tool experience, such as installation, documentation, ease of 
use, new ideas to improve the tool, etc. Appendix \ref{app:survey} has the complete survey 
with the answers.

@@

Related to the tool's functionalities, both the student assistant and the student reported 
a missing mechanism to save the tool results in a separate file instead of the terminal output. 
The tool, by default, outputs with colors, as shown in section \ref{subsec:func}, which complicates 
redirecting the terminal output to a file.

We asked for the overall impression of the tool and if it met their expectations. We also asked 
them to give a score of one to five for installation, documentation, preprocessing, and functionalities. 
Both the teaching assistant and the student scored at least four on each mentioned point, thought that 
the tool fulfilled their expectations, and had an overall good experience. Finally, the teaching assistant 
and the student reported features to the tool that may be interesting, which 
are shown below.

@@

With the survey feedback, we understood that users had a good experience and found value in the 
ArKanjo tool. They also provided critical points that could be improved. The most critical 
reported issues have been fixed, such as documentation errors and a lack of support for multiple Linux distributions 
and different operating systems. Less critical issues remain open points that 
can be addressed in future work.

@@

The second group opted to work on a webpage for the tool documentation and support for 
light and dark modes. The colored output presented in the figures in Chapter \ref{cha:tool} was 
hard-coded at first, not considering the user terminal colors, thus giving a bad user experience. 
The website helps facilitate people's contribution or use of the tool. We did not know how to 
approach the light and dark modes at first and did not need to look at how to fix the issue as 
the group searched and found a solution.

### 2025-07-24 conteudo/resumoabstract.tex
update keywords
another grammar typos fix

% As palavras-chave são obrigatórias, em português e em inglês, e devem ser
% definidas antes do resumo/abstract. Acrescente quantas forem necessárias.
\palavraschave{Linux, Duplicação de código, ferramenta}

\keywords{Linux,Code duplication,Tool}

@@

The Linux kernel is a widely used Free/Libre/Open Source Software (FLOSS) project, powering
96.3\% of the top one million web servers 
and 85\% of smartphones worldwide. Given its extensive user base, any addition of features, 
bug fixes, or handling of security vulnerabilities can impact millions. Maintaining the kernel 
is an enormous task, involving more than 28 million lines of code and contributions from over 
twenty thousand developers. Due to the scale of the project, not all contributions adhere to best 
practices, resulting in poor-quality code artifacts that may complicate maintenance and feature 
development. One such artifact is code duplication, which is the focus of this research. 
Existing tools for detecting code duplication typically compare two code artifacts to determine 
if they are duplicates. However, these tools are not well-suited for identifying duplications 
in large-scale codebases like the Linux kernel, nor do they guide on resolving detected 
duplications. Despite searches in both formal and grey literature, existing solutions fall 
short of addressing the specific challenges of the Linux kernel.
This research presents ArKanjo  , a command-line tool for Linux kernel maintenance designed to detect and
analyze function-level duplications. Released under the MIT
license, ArKanjo employs a two-stage architecture consisting
of a Preprocessor and a Query Responder that separates
computationally intensive analysis from efficient querying for duplications within large codebases.
Second, it analyzes the identified duplications to extract patterns 
and refactoring methods, offering guidance to future contributors. We evaluate ArKanjo 
against real world duplication cases in recent kernel versions, demonstrating
its effectiveness in identifying problematic clones that generic
tools often overlook. By identifying well-defined, manageable
duplication instances, ArKanjo effectively lowers the barrier
for new contributors, a capability evidenced by its role in guiding students to make their first 
code improvements to the kernel.

>>>>

% As palavras-chave são obrigatórias, em português e em inglês, e devem ser
% definidas antes do resumo/abstract. Acrescente quantas forem necessárias.
\palavraschave{Software livre, kernel Linux, duplicação de código, refatoração, manutenção de software}

\keywords{Free software,Linux kernel, Code duplication, Refactoring, code maintenance}

@@

The Linux kernel is a widely used Free/Libre/Open Source Software (FLOSS) project, powering
96.3\% of the top one million web servers 
and 85\% of smartphones worldwide. Given its extensive user base, any addition of features, 
bug fixes, or handling of security vulnerabilities can impact millions. Maintaining the kernel 
is an enormous task involving over 28 million lines of code and contributions from over 
twenty thousand developers. Due to the scale of the project, not all contributions adhere to best 
practices, resulting in poor-quality code artifacts that may complicate maintenance and feature 
development. One such artifact is code duplication, which is the focus of this research. 
Existing tools for detecting code duplication typically compare two code artifacts to determine 
if they are duplicates. However, these tools are not well-suited for identifying duplications 
in large-scale codebases like the Linux kernel, nor do they guide on resolving detected 
duplications. Despite searches in both formal and grey literature, existing solutions fall 
short of addressing the specific challenges of the Linux kernel.
This research presents ArKanjo  , a command-line tool for Linux kernel maintenance designed to detect and
analyze function-level duplications. Released under the MIT
license, ArKanjo employs a two-stage architecture consisting
of a Preprocessor and a Query Responder that separates
computationally intensive analysis from efficient querying for duplications within large codebases.
Second, it analyzes the identified duplications to extract patterns 
and refactoring methods, offering guidance to future contributors. We evaluate ArKanjo 
against real-world duplication cases in recent kernel versions, demonstrating
its effectiveness in identifying problematic clones that generic
tools often overlook. By identifying well-defined, manageable
duplication instances, ArKanjo effectively lowers the barrier
for new contributors, a capability evidenced by its role in guiding students to make their first 
code improvements to the kernel.
